---
title: "STSCI4740 Project PCR"
author: "Zhengyan Yuan"
date: "12/1/2020"
output: html_document
---

* Packages
```{r}
#install.packages('pls')
library(pls)
#install.packages("FactoMineR")
library("FactoMineR")
#install.packages("factoextra")
library("factoextra")
library(ggplot2)
```


* prepare data
```{r}
data <- read.csv("cancer_data.csv")
data <- data[,-1]
data$Level = as.factor(data$Level)
```

* principal components regression
```{r}
set.seed(1)
data$Level = as.numeric(as.factor(data$Level))
pcr.fit=pcr(Level~., data=data,scale=TRUE,validation="CV") #standardize the covariate and use CV to choose M
summary(pcr.fit)  # M=16 has smallest CV error
validationplot(pcr.fit,val.type="MSEP")

cancer.pr <- prcomp(data[c(2:23)], center = TRUE, scale = TRUE)
summary(cancer.pr)
screeplot(cancer.pr, type = "l", npcs = 15, main = "Screeplot of the first 15 PCs")
abline(h = 1, col="red", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)
cumpro <- cumsum(cancer.pr$sdev^2 / sum(cancer.pr$sdev^2))
plot(cumpro[0:15], xlab = "PC #", ylab = "Amount of explained variance", main = "Cumulative variance plot")
abline(v = 6, col="blue", lty=5)
abline(h = 0.88759, col="blue", lty=5)
legend("topleft", legend=c("Cut-off @ PC6"),
       col=c("blue"), lty=5, cex=0.6)
```
From the summary, we can tell the percentage of variance explained in the predictors and in the response using different numbers of components. When M=1, there is only 39.58% of all the variance been captured. While when M=7, it increases to 81.06.

In addition, from the screeplot of the eigenvalues and cumulative variance plot, we can see that the first 6 components has an eigenvalue >1, but until the 9th component, it explains about more than 90% of variance.

Consequently, we use M=9 component to compuet the test MSE as follows, since it explains much variance and also has relatively high eigen value. 

* Visualize the result using all features
```{r}
data$Level = as.factor(data$Level)    
fviz_pca_ind(cancer.pr, geom.ind = "point", pointshape = 21, 
             pointsize = 2, 
             fill.ind = data$Level, 
             col.ind = "black", 
             palette = "jco", 
             addEllipses = TRUE,
             label = "var",
             col.var = "black",
             repel = TRUE,
             legend.title = "Level is") +
  ggtitle("2D PCA-plot from 22 feature dataset") +
  theme(plot.title = element_text(hjust = 0.5))   
```

* Test MSE calculated
```{r}
set.seed(1)
training_index<-sample(1:nrow(data),0.8*nrow(data))
pcr.pred=predict(pcr.fit ,data[-training_index ,], ncomp =12)
data.te = data[-training_index,]
mean(as.factor(round(pcr.pred))==data.te$Level)

```

